---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gdho

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

The goal of `gdho` is to provide a detailed list of global humanitarian
organizations complied by [Humanitarian
Outcomes](https://www.humanitarianoutcomes.org/projects/gdho) which
contains basic information such as organization website and headquarter
location, as well as, operational information such as annual
expenditure. The original database is updated annually and this package
version uses data retrieved on September 20, 2023.

## Installation

You can install the development version of gdho from
[GitHub](https://github.com/gdho) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/gdho")
```

Alternatively, you can download the individual datasets as CSV or XLSX
file from the table below.

| dataset   | CSV              | XLSX              |
|-----------|------------------|-------------------|
| gdho      | [Download CSV]() | [Download XLSX]() |
| gdho_full | [Download CSV]() | [Download XLSX]() |

# Introduction

# Data

The package provides access to XX datasets ...

```{r}
library(gdho)
```

```         
The `gdho` data set has {ncol(gdho)} variables and {nrow(gdho)} observations. For an overview of the variable names, see the following table.  
```

```{r, eval=FALSE}
gdho
```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "gdho.rda") |>  ##TODO: FILL ME
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```

## Example

Write a demo snippet here.

```{r}

```

# License

Data are available as
[CC-BY](https://github.com/openwashdata/gdho/blob/main/LICENSE.md).

# Citation

```{r}
citation(gdho)
```

## Related References
