---
title: "Where are the headquarters of humanitarian organizations?"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
# read in libraries
library(gdho)
library(tidyverse)
library(sf) # for geolocation 
library(rnaturalearth) # for geolocation 
library(viridis)  
```

In this demo, we will show you how to plot a map to reflect the headquarter concentration of the humanitarian headquarters around the world. At the end of the demo, you will learn how to:

-   Summarize headquarter information from `gdho` data
-   Combine `gdho` data with external geolocation data
-   Visualize a choropleth map about headquarter location distribution

# 1. Data processing: summarize headquarters

First, we want to get a table showing how many headquarters each country has. We transform the column `hq_location` to get the information:

```{r}
gdho_hq <- gdho |>
  filter(!is.na(hq_location)) |> # remove NA data 
  group_by(hq_location) |> # group by countries
  summarise(count = n()) |> # count the number of headquarters of each country
  arrange(desc(count)) # sort by descending order
  
gdho_hq |>
  head()
```

We can see that the top six countries having the most humanitarian organization headquarters are United States, Somalia, Afgahnistan, Pakistan, Colombia and Ukraine.

# 2. Data combination: join with geodata

To plot a choropleth map, we also need to acquire geolocation information of the countries (e.g. the borders). Here we use the package `rnaturalearth` and `sf` to help. The package `rnaturalearth` contains geodata about the countries in the world, and we simply import this dataset and combine it with our `gdho` data.

```{r}
world <- rnaturalearth::ne_countries(returnclass = "sf")

ho_map <- left_join(world, gdho_hq, by = c("name_long" = "hq_location"))
```

# 3. Map visualization: concentration of the headquarters

Now, we have a data table that contains all information we need: the counts of humanitarian organization headquarters each country has, and the geo-information of each country. We start by plotting the world view of the headquarter concentration.

```{r world_map}
ggplot() +
  theme_void() +
  geom_sf(data = ho_map, aes(fill = count), color = "white", lwd = 0) + # plot map
    scale_fill_viridis(trans = "log", breaks=c(1,10,50,100,300), 
                       name = "Num. Headquarters",
                       guide = guide_legend(keyheight = unit (2, units = "mm"),
                                            label.position = "bottom", 
                                            title.position = 'top', nrow = 1)) + 
  labs(
    title = "Humanitarian Organization Headquarters Concentration",
    subtitle = "Number of headquarters per country"
  ) + # add metadata
  theme(
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.position = c(0.13, 0.15)
  ) + # polish the appearance
  coord_sf(ylim = c(-50, 90)) # Remove antarctic
```

Now we have seen how the headquarters are distributed globally, it may worth zoom in to see in detail. For example, let's check out Africa:

```{r africa_map}
af_map <- ho_map |> filter(continent == "Africa") # Filter data with only African countries
ggplot() +
  theme_void() +
  geom_sf(data = af_map, aes(fill = count), color = "white", lwd = 0.5) +
    scale_fill_viridis(trans = "log", breaks=c(1,5,10,20,50,100)) +
  labs(
    title = "Humanitarian Organization Headquarters Concentration",
    subtitle = "Number of headquarters per country in Africa"
  ) 
```

References:

[1] [Choropleth map with R and ggplot2](https://r-graph-gallery.com/327-chloropleth-map-from-geojson-with-ggplot2.html)
